Improving the Replicability of Psychological Science Through Pedagogy

### 2017-06-23 paper/paper.Rmd
mcf 23 - 33 changes

author-note:
    "*These authors contributed equally and are listed alphabetically. Thanks to the Stanford Department of Psychology and the Vice Provost for Graduate Education for funding to support the class. We are grateful to the authors of the original studies who provided materials and gave comments on a draft of this manuscript. RXDH, ENS, and MCF designed the project, supported data planning and data collection for all projects, analyzed the data, and wrote the paper. CA, JMA, RC, EH, MK, AL, SR, JR, SS, JS, and JT planned their individual projects, programmed their studies, collected their data, and analyzed their data."

@@

@@

@@

We begin by describing the general method of the class, then discuss shared methods for the replications and the process of development for each project. In general, we report how we determined our sample size, all data exclusions, all manipulations, and all measures in the study.

@@

All study participants were recruited on AMT. Individual sample sizes are given in Table \ref{tab:summary}. Each sample was recruited independently, using the same account but a different title. All replications were approved by the Stanford University Institutional Review Board under protocol \#23274, "Reproducibility of psychological science and instruction."

@@

After Pilot B review was completed, students were given authorization to collect the full sample of participants, contingent on having made any requested changes. Prior to data collection, the analytic script containing all confirmatory analyses was pre-registered using the Open Science Framework. In addition, prior to data collection for all projects, we pre-registered the confirmatory analyses reported in the current paper ([https://osf.io/rxz8m/](https://osf.io/rxz8m/)). All code and data to reproduce the analyses reported here are available at: [https://osf.io/98ta4](https://osf.io/98ta4).

@@

In this paper, we reported a series of 11 student replications of previously-published empirical studies from the 2015 volume of *Psychological Science*. We had two interlocking goals: first, to create a proof of concept for pedagogical replications, and second, to provide evidence on the ability of students to replicate recent findings in a top journal. Importantly, rather than attempting to gauge the truth of particular effects, our project aimed to assess the challenges of replicating findings -- selected on grounds of feasibility and interest -- within the constraints of a course project. Despite this narrower scope, we believe that insights from our approach are applicable to the broader goal of assessing replicability at scale.

@@

Because they were conducted by students within the constraints of a course, our studies had a number of limitations, including at least: 1) limited time for iterative piloting and adjustment, 2) limits on funding for larger samples, and 3) limits on domain expertise with respect to the specific effects that were chosen. But within these limitations, our group was able to produce a set of relatively close replications with generally good statistical power. Thus, our work here shows what is possible for motivated students using freely-available tools and resources. Such classes could become the backbone of future collaborative replication efforts [@frank2012;@everett2015].

Though others' courses that incorporate replication projects will look quite different from our own, we recommend having the replication as a centerpiece of the course and giving ample opportunity for feedback throughout the study design, as these components were crucial to our success. To encourage others to share our materials, our course outline, project templates, and assignments are available publicly at [https://osf.io/98ta4](https://osf.io/98ta4). In addition, all of the most recent lecture slides and materials are avaialable on our course website at [http://psych254.stanford.edu](). We hope that others will share and reuse our materials as they consider how best to design a course customized to their context.

@@

@@

Many statistical factors leading to lowered replicability have been discussed in past work, including analytic flexibility, context dependency, publication bias, and low statistical power among others [e.g., @button2013; @vanbavel2016; @ioannidis2005]. For both reasons of design and scale, our study here cannot disentangle these factors empirically. Nevertheless, the decreases in effect size and evidential value we report seem consistent with two explanations. First, we likely saw a "winner's curse," with initial publications tending to over-estimate effects. Second, we also might have seen a reduction in effect size because the original studies were tailored to their specific context and population, whereas our replications may not have been [@vanbavel2016, but cf. @inbar2016]. Future work would benefit from a statement of constraints on generality so as to provide a guide to the conditions under which an effect is likely to be present [@simons2016]. Both of these explanations are consistent with our observations, and we suspect both winner's curse and context effects are present to some degree.

@@

Indeed, no study we attempted to replicate was pre-registered (though some were internal replications), and the general pattern of evidence across studies was often more important to the authors' conclusions than any particular test in any single study.  Almost every study conducted multiple statistical tests, often associated with several statistical models with differing specifications. And in some cases, there was no conventionally-reported and easily calculable effect size measures [e.g., for mixed- and random-effect models, cf. @bakeman2005]. This set of features -- which, in our experience, are endemic to published psychological work, including our own -- makes replication research within the conventional statistical paradigm extremely challenging.

In our view, two recommendations will improve this situation dramatically. First, authors should specify their "key test" more clearly. Pre-registration might facilitate this step, as regstration typically requires the selection of particular statistical analyses to be designated as critical for interpretation, but we believe this designation is critical even in un-registered experiments. Had "key test" information been available, it would have dramatically simplified the problem of selecting the effect to replicate in several cases. Second, the widespread adoption of open data practices would dramatically facilitate the type of meta-research pursued here. In addition to clarifying the (often terse) reporting in a paper, the availability of data facilitates the computation of the relevant meta-analytic variables.^[We note that, although we were grateful that several authors of the papers in our sample shared data and materials, the use of open (non-proprietary) formats for archiving data -- and in particular analytic code -- is often critical for investigators attempting to reproduce analyses.] No paper can perfectly predict the uses to which future investigators will put a piece of work; the availability of the source materials is thus an important route for more flexible reuse.

>>>>

author-note:
    "*These authors contributed equally and are listed alphabetically."

@@

To encourage others to share our materials, our course outline, project templates, and assignments are available publicly at [https://osf.io/98ta4](https://osf.io/98ta4). In addition, all of the most recent lecture slides and materials are avaialable on our course website at [http://psych254.stanford.edu](). We hope that others will share and reuse our materials as they consider how best to design a course customized to their context.

@@

# Disclosures

## Preregistration

<!-- Preregistration: Provide a link to any preregistration documentation. If only a subset of reported studies were preregistered, indicate which ones were and which ones were not. -->

As described below, our procedure for individual project preregistrations was to register the analytic script with specific key hypothesis tests clearly marked. Individual project preregistration links are given in Table 1. Prior to data collection for all projects, we also pre-registered the confirmatory analyses reported in the current paper ([https://osf.io/rxz8m/](https://osf.io/rxz8m/)).

## Data, Materials, and Online Resources

<!-- Data, materials, and online resources: Provide a permanent, persisting link to a public archive where readers can access any code, materials, de-identified data, and other resources (e.g., osf.io, perma.cc, clinicaltrials.gov). If any materials or data are not publicly available, the article should explain why and should note how researchers can access them for research purposes. -->

All code and data necessary to reproduce the analyses reported here are available at: [https://osf.io/98ta4](https://osf.io/98ta4).

## Measures

<!-- Measures: The text here should state “We report how we determined our sample size, all data exclusions, all manipulations, and all measures in the study” (see Simmons, Nelson, & Simonsohn, 2011). If any aspect of this statement is untrue, authors should explain why. All such details should be specified in the preregistration documentation as well. -->

We report how we determined our sample size, all data exclusions, all manipulations, and all measures in the study (see @simmons2011).

## Subjects

<!-- Subjects: If the article reports research with human or animal subjects, authors should state that testing was conducted with the approval of an institutional review board or similar ethics committee (with the protocol number provided). For those involving human subjects, authors should state whether the research was conducted in accordance with the Declaration of Helsinki; note that the latest version of the Declaration of Helsinki requires preregistration before data collection begins. -->

All replications were approved by the Stanford University Institutional Review Board under protocol \#23274, "Reproducibility of psychological science and instruction" and was conducted in accordance with the Declaration of Helsinki.

## Conflicts of Interest

<!-- Conflicts of Interest: Authors should identify any conflicts of interest here and should also report them during the submission process. If authors have no conflicts of interest, they should state, “” -->

The authors declare that they have no conflicts of interest with respect to the authorship or the publication of this article.

## Author Contributions

<!-- Author Contributions: Authorship implies significant participation in the research reported or in writing the manuscript, including participation in the design and/or interpretation of reported experiments or results, participation in the acquisition and/or analysis of data, and participation in drafting and/or revising the manuscript. All authors must agree to the order in which the authors are listed and must have read the final manuscript and approved its submission. They must also agree to take responsibility for the work in the event that its integrity or veracity is questioned. Authors should precisely describe their contributions to the research and manuscript, identifying each author by his or her initials. For example, “DJS and AOH jointly generated the idea for the study. AOH programmed the study and collected the data. DJS wrote the analysis code and analyzed the data, and AOH verified the accuracy of those analyses. DJS wrote the first draft of the manuscript, and both authors critically edited it. Both authors approved the final submitted version of the manuscript.” -->

RXDH, ENS, and MCF designed the project, supported data planning and data collection for all projects, analyzed the data, and wrote the paper. CA, JMA, RC, EH, MK, AL, SR, JR, SS, JS, and JT planned their individual projects, programmed their studies, collected their data, and analyzed their data.

## Acknowledgements

<!-- Acknowledgments: Authors should use this section to identify any people who should be credited for their assistance with the reported research. -->

Thanks to the Stanford Department of Psychology and the Vice Provost for Graduate Education for funding to support the class. We are grateful to the authors of the original studies who provided materials and gave comments on a draft of this manuscript.

## Prior Versions

<!-- Prior versions: If part or all of a submitted manuscript was previously posted to a blog or to a preprint archive, provide a link to those sources and briefly indicate what aspects of the manuscript are shared between the submitted version and earlier versions. -->

An earlier draft of this manuscript was posted at [https://osf.io/preprints/psyarxiv/p73he/](https://osf.io/preprints/psyarxiv/p73he/).

@@

We begin by describing the general method of the class, then discuss shared methods for the replications and the process of development for each project.

@@

All study participants were recruited on AMT. Individual sample sizes are given in Table \ref{tab:summary}. Each sample was recruited independently, using the same account but a different title.

@@

After Pilot B review was completed, students were given authorization to collect the full sample of participants, contingent on having made any requested changes. Prior to data collection, the analytic script containing all confirmatory analyses was pre-registered using the Open Science Framework.

@@

In this paper, we reported a series of 11 student replications of previously-published empirical studies from the 2015 volume of *Psychological Science*. Our goal was to provide a proof of concept for pedagogical replications of recent findings in a top journal. Importantly, rather than attempting to gauge the truth of particular effects, our project aimed to assess the challenges of replicating findings -- selected on grounds of feasibility and interest -- within the constraints of a course project. Despite this narrower scope, we believe that insights from our approach are applicable to the broader goal of assessing replicability at scale, with implications both for replication pedagogy and for the practice of replication more broadly. We address each in turn.

@@

Because they were conducted by students within the constraints of a course, our studies had a number of limitations, including at least: 1) limited time for iterative piloting and adjustment, 2) limits on funding for larger samples, and 3) limits on domain expertise with respect to the specific effects that were chosen. But within these limitations, our group was able to produce a set of relatively close replications with generally good statistical power. Thus, our work here shows what is possible for motivated students using freely-available tools and resources.

Such classes could become the backbone of future collaborative replication efforts [@frank2012;@everett2015]. This kind of cross-class collaboration would require instructors to make class results more broadly discoverable. We can envision a number of possible avenues for this kind of sharing. For example, if some products are shared openly in a repository like the [Open Science Framework](http://osf.io) or [figshare](http://figshare.com) and tagged appropriately, they will be discoverable via search engines. Alternatively, a more structured method for sharing and discovery would be upload to specific replication curation websites like [PsychFileDrawer](http://psychfiledrawer.org) or CurateScience (http://curatescience.org/). Both of these models assume limited coordination across instructors, but more structured collaborations are of course possible. For example, the [Collaborative Replications and Education Project](https://osf.io/wfc6u/), selected a subsample of important and feasible studies for replication and helped provide materials to instructors with the explicit goal of encouraging cross-lab pedagogical replications.

One challenge for this kind of dissemination is how to curate which projects are released publicly. We are generally optimistic about the contributions of student work, but of course not every student project will be scientifically informative. To take an edge case, imagine a student makes an error in stimulus randomization that is not caught by the instructors until after data collection and that and creates a major confound in the collected data. In this case, we would argue that this work should not be deposited in a targeted replication repository. In the end, these decisions will be the responsibility of the course instructor -- much like decisions about when to publish other research.

Though others' courses that incorporate replication projects will look quite different from our own, we recommend having the replication as a centerpiece of the course and giving ample opportunity for feedback throughout the study design. Without a central focus on the replication project, students would not have the time or motivation to complete a high quality project; without multiple opportunities for instructor feedback, we would have far less confidence in the quality of the final projects.

@@

## Decision Points for Instructors

- how to customize for different levels of instructor

### Ethical approvals

### Study selection

- student expertise
- group wiork
- Strength of evidence in original, e.g. BF > 3

### Contacting original authors

### Results deposit

@@

Many statistical factors leading to lowered replicability have been discussed in past work, including analytic flexibility, context dependency, publication bias, and low statistical power among others [e.g., @button2013; @vanbavel2016; @ioannidis2005]. For both reasons of design and scale, our study here cannot disentangle these factors empirically. Nevertheless, the decreases in effect size and evidential value we report seem consistent with two explanations. First, we likely saw a "winner's curse," with initial publications tending to over-estimate effects. Second, we also might have seen a reduction in effect size because the original studies were tailored to their specific context and population, whereas our replications may not have been [@vanbavel2016, but cf. @inbar2016]. Future work would benefit from a statement of constraints on generality so as to provide a guide to the conditions under which an effect is likely to be present [@simons2016]. Both of these explanations are consistent with our observations, and, based on the previous literature [@osf2015,@vanbavel2016], we speculate that both winner's curse and context effects were present to some degree.

@@

Indeed, no study we attempted to replicate was pre-registered (though some were internal replications), and the general pattern of evidence across studies was often more important to the authors' conclusions than any particular test in any single study.  Almost every study conducted multiple statistical tests, often associated with several tersely-reported statistical models with differing specifications. And in some cases, there was no conventionally-reported and easily calculable effect size measures [e.g., for mixed- and random-effect models, cf. @bakeman2005]. This set of features -- which, in our experience, are endemic to published psychological work, including our own -- makes it extremely challenging to do replication research focused on the estimation of a single effect size.

In our view, two recommendations will improve this situation. First, authors should specify their "key test" more clearly. Pre-registration might facilitate this step, as regstration typically requires the selection of particular statistical analyses to be designated as critical for interpretation, but we believe this designation is critical even in un-registered experiments. Had "key test" information been available, it would have dramatically simplified the problem of selecting the effect to replicate in several cases. Second, the widespread adoption of open data and open analytic code practices would facilitate replication research. In addition to clarifying the summary report of results in a paper, the availability of data dramatically facilitates the computation of the relevant meta-analytic effect size for power analysis and statistical comparison.^[We note that, although we were grateful that several authors of the papers in our sample shared data and materials, the use of open (non-proprietary) formats for archiving data -- and in particular analytic code -- is often critical for investigators attempting to reproduce analyses.]
